The problem and promise of scale in multilayer animal social networks

### 2020-02-15 paper/manuscript.Rmd
reflow

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods [e.g. @Carter_2015], whereas social association
represents shared space use by members of the same social group [Figure 1,
@Franks_2009]. <!--Figure \@ref(fig:spacetime) TODO: something broken here?? -->
Further, for social interactions or associations to occur individuals must share
space, and thus have overlapping home ranges [@VanderWal_2014]. In the context
of the social environment, home range overlap is an example of how animals share
space over coarser spatial and temporal scales (e.g. Piza-Roca et al. 2018).
Although the social environment clearly scales spatially and temporally, it
remains unclear whether coarser scales of sociality, including social
association and home range overlap, vary predictably with social interaction
[@Castles_2014;
@Farine_2015a]. Thus, we define social scale as the type of social relationship
observed that can be explicitly defined and measured [@Farine_2015a]. <!--
Farine references comes out as D.R. Farine 2015 -->

@@

### Temporal scale
#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies.We
investigated the influence of the number of observations on resulting multilayer
networks. We randomly selected a maximum 485 timesteps for each season to
represent all possible observations. We regenerated multilayer networks and
calculated associations of individuals with 25 additional observations at each
iteration (10 - 485 observations). Subsequent iterations included the previous set of observations to
mimic collection of observational data. <!-- TODO: bah conclude / this sentence
falls off a cliff -->

@@

# Scale in multilayer networks in practice: case study results and discussion
<!-- TODO: Add intro/general to results? with mention figures are means of inds-->

@@

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015] With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues than in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

@@

Social processes, movement decisions, and space use are inextricably linked and
multilayer networks represent a potential analytical framework for linking these
processes [@Mourier_2019]. Movement as a collective group
can reduce per capita predation risk via detection-dilution trade-offs and
improve information transfer about the quality or location of resources to other
group members [@Jolles_2019]. In our case study, we integrate proximity based
social networks that inherently rely on animal collective movement and, in
general, we found that movement is likely an important mechanism driving social
network connectivity within and between habitat types.

>>>>

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods [e.g. @Carter_2015], whereas social association
represents shared space use by members of the same social group [Figure
\@ref(fig:spacetime), @Franks_2009]. <!--Figure \@ref(fig:spacetime) TODO:
something broken here?? --> Further, for social interactions or associations to
occur individuals must share space, and thus have overlapping home ranges
[@VanderWal_2014]. In the context of the social environment, home range overlap
is an example of how animals share space over coarser spatial and temporal
scales (e.g. Piza-Roca et al. 2018). Although the social environment clearly
scales spatially and temporally, it remains unclear whether coarser scales of
sociality, including social association and home range overlap, vary predictably
with social interaction [@Castles_2014;
@Farine_2015a]. Thus, we define social scale as the type of social relationship
observed that can be explicitly defined and measured [@Farine_2015a]. <!--
Farine references comes out as D.R. Farine 2015 -->

@@

### Temporal scale
#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies.We
investigated the influence of the number of observations on resulting multilayer
networks. We randomly selected a maximum 485 timesteps for each season to
represent all possible observations. We regenerated multilayer networks and
calculated associations of individuals with 25 additional observations at each
iteration (10 - 485 observations). Subsequent iterations included the previous
set of observations to mimic collection of observational data.

@@

# Scale in multilayer networks in practice: case study results and discussion
<!-- TODO: Add intro/general to results?-->

@@

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues unlike in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

@@

Social processes, movement decisions, and space use are inextricably linked and
multilayer networks represent a potential analytical framework for linking these
processes [@Mourier_2019]. Movement as a collective group
can reduce per capita predation risk via detection-dilution trade-offs and
improve information transfer about the quality or location of resources to other
group members [@Jolles_2019]. In our case study, we integrate proximity based
social networks that inherently rely on animal collective movement, and, in
general, we found that movement is likely an important mechanism driving social
network connectivity within and between habitat types.
